TacticToe: Learning the Game of Tactics

### 2018-02-07 tactictoe_jv.tex
Move table of contents before content

\section{Introduction}
\tableofcontents

\todoi{Comparison with hammers in usability}
Many of the state-of-the-art interactive theorem provers (ITPs) such as
  \holfour~\cite{hol4}, \hollight~\cite{Harrison09hollight},
  \isabelle~\cite{isabelle}
  and \coq~\cite{coq-book}
  provide high-level parameterizable tactics for constructing the
  proofs.  Such tactics typically analyze the current goal state and
  assumptions, apply nontrivial proof transformations, which get
  expanded into possibly many basic kernel-level inferences or significant
  parts of the proof term.
%\marginpar{not Coq}
  In this
  work we develop a tactic-level automation procedure for the \holfour ITP
  which guides selection of the tactics by learning from previous
  proofs.  Instead of relying on translation to first-order automated
  theorem provers (ATPs) as done by the hammer
  systems~\cite{hammers4qed,tgck-cpp15}, the technique
%\marginpar{too many cites?}
%BlanchetteGKKU16,holyhammer,mizAR40
  directly searches for sequences of tactic applications that lead to
  the ITP proof, thus avoiding the translation and
  proof-reconstruction phases needed by the hammers.

\todoi{Rewrite the plan}
  To do this, we \emph{extract and record} tactic invocations from the ITP
  proofs (Section~\ref{sec:recording}) and
  \emph{build efficient machine learning classifiers} based on such training
  examples (Section~\ref{s:prediction}).  The learned data serves as a
  guidance for a \emph{Monte Carlo tree search} that explores the different
  proof paths (Section~\ref{sec:proofsearch}). The result, if
  successful, is a certified human-level proof composed of \holfour
  tactics.  The system is evaluated on a large set of theorems originating
  from \holfour (Section~\ref{s:experiments}), and we show that the performance
  of the single best \tactictoe strategy exceeds the performance of a hammer
  system used with a single strategy and a single efficient external prover.

>>>>

\tableofcontents
\todoi{Comparison with hammers in usability}

\section{Introduction}

Many of the state-of-the-art interactive theorem provers (ITPs) such as
  \holfour~\cite{hol4}, \hollight~\cite{Harrison09hollight},
  \isabelle~\cite{isabelle}
  and \coq~\cite{coq-book} provide high-level parameterizable tactics for constructing proofs.
  Such tactics typically analyze the current proof state (goal and
  assumptions), then apply non-trivial proof transformations that get
  expanded into many basic kernel-level inferences or large parts of a proof term.
%\marginpar{not Coq}
  In this work we develop a tactic-level automation procedure for the \holfour ITP
  which guides the selection of tactics by learning from previous
  proofs.  Instead of relying on translation to first-order automated
  theorem provers (ATPs) as done by the hammer
  systems~\cite{hammers4qed,tgck-cpp15}, our technique
%\marginpar{too many cites?}
%BlanchetteGKKU16,holyhammer,mizAR40
  searches directly for sequences of tactic applications that lead to
  an ITP proof, thus avoiding the translation and proof-reconstruction phases needed by the hammers.

\todoi{Rewrite the plan}
  To do this, we \emph{extract and record} tactic invocations from the ITP
  proofs (Section~\ref{sec:recording}) and
  \emph{build efficient machine learning classifiers} based on such training
  examples (Section~\ref{s:prediction}).  The learned data serves as
  guidance for a \emph{Monte Carlo tree search} that explores different
  proof paths (Section~\ref{sec:proofsearch}). The result, if
  successful, is a certified human-level proof composed of \holfour
  tactics.  The system is evaluated on a large set of theorems originating
  from \holfour (Section~\ref{s:experiments}), and we show that the performance
  of the single best \tactictoe strategy exceeds the performance of a hammer
  system used with a single strategy and a single efficient external prover.
